---
title: "Selection pressures in _Syngnathus scovelli_"
output: 
  github_document:
    toc: true
header-includes: >
  \usepackage{lipsum}
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir='../',
                     fig_path="../figs/")
```

``` {r library, message = FALSE, warning = FALSE}
#This is a cohesive list of all the libraries used in this document
library(ggplot2)
library(cowplot)
library(fBasics)
library(pwr)
library(dplyr)
library(tidyr)
library(knitr)

```

``` {r read-data, message = FALSE, warning = FALSE}

#Metadata for males and females from the mesocosm experiments
fem_mesoSS <- read.csv("data/all_fem_meso_scovelli.csv")
mal_mesoSS <- read.csv("data/all_mal_meso_scovelli.csv")

```

This document will follow the same analysis as was outlined in `selection_analysis_floridae.Rmd`. For more thorough details refer back to that document. 

The datasets used in this document were pulled from a publically accessible manuscript that ran similar experimental breeding populations (Rose et al., 2013).

# Cleaning the datasets
For the male dataset I want to do a few things before using it: 
  1. Remove the two males that died (C1M5 and C6M2).
  2. Replace all of the "NAs" present in the males who didn't mate with 0's when appropriate.
  3. Add a column of mating success to the dataset. This will be either a 0 or 1 as _S. scovelli_ males only mate once.

In the female dataset I am going to remove C1F2 since she doesn't have any data related to her reproductive success.

For both the male and the female datasets I also want to subset them out to only include the control fish and not fish that were exposed to estrogen.

```{r clean-datasets}
#Subset the datasets to remove the fish exposed to estrogen
fem_succSS <- fem_mesoSS[grep("C", fem_mesoSS$trial_num),]
mal_succSS <- mal_mesoSS[grep("C", fem_mesoSS$trial_num),]

#Adding full fishIDs to make removing individuals easier
fem_succSS$femID <- paste0(fem_succSS$trial_num, "F",
                         fem_succSS$fishID)
mal_succSS$maleID <- paste0(mal_succSS$trial_num, "M",
                          mal_succSS$fishID)
#Removing the one female
fem_succSS <- subset(fem_succSS, !(femID %in% "C1F2"))

#Removing the two males who died
mal_succSS <- subset(mal_succSS, !(maleID %in% c("C1M5", "C6M2")))

#Replace NAs with 0s in the columns related to fitness
mal_succSS[, c("NumDeveloped", 
               "NumUndeveloped",
               "totalEggs")] <- sapply(mal_succSS[, c("NumDeveloped",
                                                      "NumUndeveloped",
                                                      "totalEggs")],
                                       function(x)
                                         as.numeric(ifelse(is.na(x), 0, x)))

#Add a column for females and males to denote mating success
mal_succSS$MatingSuccess <- ifelse(mal_succSS$totalEggs > 0, 
                                   1, 
                                   0)

fem_succSS$mated <- ifelse(fem_succSS$MatingSuccess > 0,
                           1,
                           0)

```

# Calculating the degree of sexual dimorphism
The original datasets contained only information about standard length. To allow for a more complete comparison between all three species within the genus _Syngnathus_ I obtained the original photographs of those pipefish and re-measured them to obtain measurements of both torso depth and snout-vent length for males and females. For males I also generated additional measurements of pouch area and pouch length.

Now, we can see if males and females differ in standard length, torso depth, and snout-vent length. First, I need to see if assumptions are met, i.e. variances are equal and data is normally distributed.

## Checking the assumptions for a pairwise comparison
 The main two things that I will be looking into include:

1. Equal variances between groups (using `var.test()`).
2. Normal distribution of the data (using `normalTest()`).

To account for the fact that fish who are longer may just inherently be deeper as well, I am going to adjust the depth by the snout-vent length of the pipefish prior to running any analyses. This svl-adjusted depth will then be used in all consequent analyses.

```{r sd-assumption-test}

#Adjust the torso depth
fem_succSS$depth_adj <- fem_succSS$depth/fem_succSS$svl
mal_succSS$depth_adj <- mal_succSS$depth/mal_succSS$svl

#Testing to see if the variances are equal
var.test(fem_succSS$length, mal_succSS$length) #not equal
var.test(fem_succSS$depth_adj, mal_succSS$depth_adj) #not equal
var.test(fem_succSS$svl, mal_succSS$svl) #equal

#Testing for normal distribution
normalTest(fem_succSS$length, method = "da") #not normal
normalTest(mal_succSS$length, method = "da") #normal
normalTest(fem_succSS$depth_adj, method = "da") #normal
normalTest(mal_succSS$depth_adj, method = "da") #normal
normalTest(fem_succSS$svl, method = "da") #not normal
normalTest(mal_succSS$svl, method = "da") #normal

```

## Investigate distributions and run the tests
I will run a Wilcoxon test for standard length and snout-vent length and a Welch's two-sample t-test for depth.

```{r histogram_sizes, message=FALSE, warning = FALSE, echo = FALSE, fig.cap="_Histograms of male and female pipefish body sizes._", fig.width=15}

#Combining the datasets
fem_succSS$Sex <- "F"
mal_succSS$Sex <- "M"
all_sex_meso <- rbind(fem_succSS[, c("length", "depth_adj", "svl", "Sex")], mal_succSS[, c("length", "depth_adj", "svl", "Sex")])

length <- ggplot(data = all_sex_meso, aes(x = length)) +
  geom_histogram(aes(fill = Sex),
                 alpha = 0.5,
                 position = 'identity') +
  scale_fill_manual(values = c("#7fc97f75", "#beaed475")) +
  theme_minimal() +
  labs(x = "Length (mm)",
       y = "Number of Pipefish",
       title = "Length of male and female pipefish")

depth <- ggplot(data = all_sex_meso, aes(x = depth_adj)) +
  geom_histogram(aes(fill = Sex),
                 alpha = 0.5,
                 position = 'identity') +
  scale_fill_manual(values = c("#7fc97f75", "#beaed475")) +
  theme_minimal() +
  labs(x = "Depth (mm)",
       y = "Number of Pipefish",
       title = "Depth of male and female pipefish")

svl <- ggplot(data = all_sex_meso, aes(x = svl)) +
  geom_histogram(aes(fill = Sex),
                 alpha = 0.5,
                 position = 'identity') +
  scale_fill_manual(values = c("#7fc97f75", "#beaed475")) +
  theme_minimal() +
  labs(x = "SVL (mm)",
       y = "Number of Pipefish",
       title = "Snout-vent length of male and female pipefish")

plot_grid(length, depth, svl,
          ncol = 3)

```

```{r sd-t-tests}
wilcox.test(fem_succSS$length, mal_succSS$length) #Sig. difference
wilcox.test(fem_succSS$svl, mal_succSS$svl) #Sig. difference
t.test(fem_succSS$depth_adj, mal_succSS$depth_adj, 
       var.equal = FALSE) #Sig. difference

```

For the Gulf pipefish, there are significant differences between males and females in terms of standard length, snout-vent length, and torso depth. 

```{r power-test}

#Checking the power - length
d_mean_len <- abs(mean(fem_succSS$length, na.rm = TRUE) - 
                    mean(mal_succSS$length, na.rm = TRUE))
pool_sd_len <- sqrt((var(fem_succSS$length, na.rm = TRUE) + 
                       var(mal_succSS$length, na.rm = TRUE))/ 2)
d_len <- d_mean_len/pool_sd_len

pwr.t.test(n = nrow(fem_succSS), 
           d = d_len,
           sig.level = 0.05,
           type = 'two.sample',
           alternative = 'two.sided')

#Checking the power - SVL
d_mean_svl <- abs(mean(fem_succSS$svl, na.rm = TRUE) - 
                    mean(mal_succSS$svl, na.rm = TRUE))
pool_sd_svl <- sqrt((var(fem_succSS$svl, na.rm = TRUE) + 
                       var(mal_succSS$svl, na.rm = TRUE))/ 2)
d_svl <- d_mean_svl/pool_sd_svl

pwr.t.test(n = nrow(fem_succSS), 
           d = d_svl,
           sig.level = 0.05,
           type = 'two.sample',
           alternative = 'two.sided')

#Checking the power - Depth
d_mean_depth <- abs(mean(fem_succSS$depth_adj, na.rm = TRUE) - 
                      mean(mal_succSS$depth_adj, na.rm = TRUE))
pool_sd_depth <- sqrt((var(fem_succSS$depth_adj, na.rm = TRUE) + 
                         var(mal_succSS$depth_adj, na.rm = TRUE))/ 2)
d_depth <- d_mean_depth/pool_sd_depth
pwr.t.test(n = nrow(fem_succSS), 
           d = d_depth,
           sig.level = 0.05,
           type = 'two.sample',
           alternative = 'two.sided')

```

For all variables we have a power of over 0.9 or over 90% so we can be confident in our interpretation.

# Summary statistics for successfully mated individuals
## Males
Across all `r length(unique(mal_succSS$trial_num))` trials and `r nrow(mal_succSS)` total males, there were `r nrow(mal_succSS[mal_succSS$MatingSuccess == 1,])` males that mated at least one time and `r nrow(mal_succSS[mal_succSS$MatingSuccess > 1,])` of those males had two mates. 

Looking across all males, including the ones that did not mate, this is what we find as the mean, sd, and se for the number of embryos transferred and how many of those developed versus didn't:

|                  | mean| SD| SE| max | min |
|:-----------------|----:|---:|----:|----:|----:|
|Number of Embryos | `r mean(mal_succSS$totalEggs)`|`r sd(mal_succSS$totalEggs)`|  `r sd(mal_succSS$totalEggs)/sqrt(nrow(mal_succSS))`|`r max(mal_succSS$totalEggs)` | `r min(mal_succSS$totalEggs)`|
|Developed Embryos | `r mean(mal_succSS$NumDeveloped)`|`r sd(mal_succSS$NumDeveloped)`|  `r sd(mal_succSS$NumDeveloped)/sqrt(nrow(mal_succSS))`|`r max(mal_succSS$NumDeveloped)` | `r min(mal_succSS$NumDeveloped)`|
|Undeveloped Embryos| `r mean(mal_succSS$NumUndeveloped)`|`r sd(mal_succSS$NumUndeveloped)`|  `r sd(mal_succSS$NumUndeveloped)/sqrt(nrow(mal_succSS))`|`r max(mal_succSS$NumUndeveloped)` | `r min(mal_succSS$NumUndeveloped)`|

These values will be influenced by the number of 0s coming from males who did not mate. So let's look at the same thing, but this time for only males who had at least one successful mating:

|                  | mean| SD| SE|  max | min |
|:-----------------|----:|---:|----:|----:|----:|
|Number of Embryos | `r mean(mal_succSS$totalEggs[mal_succSS$MatingSuccess == 1])`|`r sd(mal_succSS$totalEggs[mal_succSS$MatingSuccess == 1])`|  `r sd(mal_succSS$totalEggs[mal_succSS$MatingSuccess == 1])/sqrt(nrow(mal_succSS[mal_succSS$MatingSuccess == 1,]))`| `r max(mal_succSS$totalEggs[mal_succSS$MatingSuccess == 1])` | `r min(mal_succSS$totalEggs[mal_succSS$MatingSuccess == 1])` |
|Developed Embryos | `r mean(mal_succSS$NumDeveloped[mal_succSS$MatingSuccess == 1])`|`r sd(mal_succSS$NumDeveloped[mal_succSS$MatingSuccess == 1])`|  `r sd(mal_succSS$NumDeveloped[mal_succSS$MatingSuccess == 1])/sqrt(nrow(mal_succSS[mal_succSS$MatingSuccess == 1,]))`|`r max(mal_succSS$NumDeveloped[mal_succSS$MatingSuccess == 1])` | `r min(mal_succSS$NumDeveloped[mal_succSS$MatingSuccess == 1])` |
|Undeveloped Embryos| `r mean(mal_succSS$NumUndeveloped[mal_succSS$MatingSuccess == 1])`|`r sd(mal_succSS$NumUndeveloped[mal_succSS$MatingSuccess == 1])`|  `r sd(mal_succSS$NumUndeveloped[mal_succSS$MatingSuccess == 1])/sqrt(nrow(mal_succSS[mal_succSS$MatingSuccess == 1,]))`| `r max(mal_succSS$NumUndeveloped[mal_succSS$MatingSuccess == 1])` | `r min(mal_succSS$NumUndeveloped[mal_succSS$MatingSuccess == 1])` |

I know want to see if there are correlations between morphometrics such as standard length and the size of the brood pouch and brood size (i.e., are larger males securing larger broods).

```{r em-v-bp, echo=FALSE, fig.cap="_Scatterplot of the relationship between brood pouch size metrics and the number of embryos a male had._", fig.width=15, fig.height=10 }

#create a subset of only the males who mated
mated_malSS <- mal_succSS[mal_succSS$MatingSuccess == 1, ]

#Plot brood size against the different metrics for pouch size
par(mfrow=c(1,2))
plot(mated_malSS$bp_area,
     mated_malSS$totalEggs,
     xlab = expression(paste("Brood Pouch Area (mm"^2*")")),
     ylab = "Brood Size (# embryos)",
     col = "darkorange",
     pch = 19)
abline(lm(mated_malSS$totalEggs ~ as.numeric(mated_malSS$bp_area)), lwd = 3, lty = 2)
plot(mated_malSS$bp_length,
     mated_malSS$totalEggs,
     xlab = "Brood Pouch Length (mm)",
     ylab = "Brood Size (# embryos)",
     col = "cyan4",
     pch = 19)
abline(lm(mated_malSS$totalEggs ~ as.numeric(mated_malSS$bp_length)), lwd = 3, lty = 2)
```

There may be some correlation happening here, but it doesn't look particularly strong. Let's run some correlations tests to see what they say.

```{r eggs-bp-cor, echo=FALSE}
cor.test(as.numeric(mated_malSS$bp_area), mated_malSS$totalEggs)
cor.test(as.numeric(mated_malSS$bp_length), mated_malSS$totalEggs)
```

There is no significant correlation with either of the brood pouch size metrics. Let's see if there is some relationship with just the overall size of the male.

```{r em-v-sl, echo=FALSE, fig.cap="_Scatterplot of the relationship between standard length (mm) and the number of embryos a male had._", fig.width=15, fig.height=10}

mal_succSS$totalEggs <- as.numeric(mal_succSS$totalEggs)

plot(mated_malSS$length,
     mated_malSS$totalEggs,
     xlab = "Standard Length (mm)",
     ylab = "Brood Size (# embryos)",
     col = "#CD5C5C75",
     pch = 19,
     cex = 1.5
)
abline(lm(mated_malSS$totalEggs ~ as.numeric(mated_malSS$length)), lwd = 3, lty = 2)
```

```{r eggs-sl-cor, echo=FALSE}
cor.test(mated_malSS$length, mated_malSS$totalEggs)
```

There is a significant correlation! Larger males have bigger broods in terms of standard length.

## Females
Across all `r length(unique(fem_succSS$trial_num))` trials and `r nrow(fem_succSS)` total females, there were `r nrow(fem_succSS[fem_succSS$MatingSuccess != 0,])` females that mated at least one time, `r nrow(fem_succSS[fem_succSS$MatingSuccess == 2,])` females that mated twice, and `r nrow(fem_succSS[fem_succSS$MatingSuccess == 3,])` that mated 3 times. 

Looking across all females, including the ones that did not mate, this is what we find as the mean, sd, and se for the total number of embryos transferred from each female (across all of her mates if applicable) and how many of those developed versus didn't:

|                  | mean| SD| SE| max | min |
|:-----------------|----:|---:|----:|----:|----:|
|Number of Embryos | `r mean(fem_succSS$totalEggs)`|`r sd(fem_succSS$totalEggs)`|  `r sd(fem_succSS$totalEggs)/sqrt(nrow(fem_succSS))`|`r max(fem_succSS$totalEggs)` | `r min(fem_succSS$totalEggs)`|
|Developed Embryos | `r mean(fem_succSS$NumDeveloped)`|`r sd(fem_succSS$NumDeveloped)`|  `r sd(fem_succSS$NumDeveloped)/sqrt(nrow(fem_succSS))`|`r max(fem_succSS$NumDeveloped)` | `r min(fem_succSS$NumDeveloped)`|
|Undeveloped Embryos| `r mean(fem_succSS$NumUndeveloped)`|`r sd(fem_succSS$NumUndeveloped)`|  `r sd(fem_succSS$NumUndeveloped)/sqrt(nrow(fem_succSS))`|`r max(fem_succSS$NumUndeveloped)` | `r min(fem_succSS$NumUndeveloped)`|

These values will be influenced by the number of 0s coming from females who did not mate. So let's look at the same thing, but this time for only females who had at least one successful mating:

|                  | mean| SD| SE|  max | min |
|:-----------------|----:|---:|----:|----:|----:|
|Number of Embryos | `r mean(fem_succSS$totalEggs[fem_succSS$MatingSuccess != 0])`|`r sd(fem_succSS$totalEggs[fem_succSS$MatingSuccess != 0])`|  `r sd(fem_succSS$totalEggs[fem_succSS$MatingSuccess != 0])/sqrt(nrow(fem_succSS[fem_succSS$MatingSuccess != 0,]))`| `r max(fem_succSS$totalEggs[fem_succSS$MatingSuccess != 0])` | `r min(fem_succSS$totalEggs[fem_succSS$MatingSuccess != 0])` |
|Developed Embryos | `r mean(fem_succSS$NumDeveloped[fem_succSS$MatingSuccess != 0])`|`r sd(fem_succSS$NumDeveloped[fem_succSS$MatingSuccess != 0])`|  `r sd(fem_succSS$NumDeveloped[fem_succSS$MatingSuccess != 0])/sqrt(nrow(fem_succSS[fem_succSS$MatingSuccess != 0,]))`|`r max(fem_succSS$NumDeveloped[fem_succSS$MatingSuccess != 0])` | `r min(fem_succSS$NumDeveloped[fem_succSS$MatingSuccess != 0])` |
|Undeveloped Embryos| `r mean(fem_succSS$NumUndeveloped[fem_succSS$MatingSuccess != 0])`|`r sd(fem_succSS$NumUndeveloped[fem_succSS$MatingSuccess != 0])`|  `r sd(fem_succSS$NumUndeveloped[fem_succSS$MatingSuccess != 0])/sqrt(nrow(fem_succSS[fem_succSS$MatingSuccess != 0,]))`| `r max(fem_succSS$NumUndeveloped[fem_succSS$MatingSuccess != 0])` | `r min(fem_succSS$NumUndeveloped[fem_succSS$MatingSuccess != 0])` |

I want to see what relationship there may be between female body size (in terms of standard length, depth, and SVL) and the number of eggs she transferred. I also want to see on average how many eggs were transferred per mating. I'm going to calculate this by taking the total number of eggs and dividing it by the number of mates. 

```{r egg-per-mate-fem, echo=FALSE}
eggs_per_mate <- fem_succSS$totalEggs/fem_succSS$MatingSuccess
mean(eggs_per_mate, na.rm = TRUE)
sd(eggs_per_mate, na.rm = TRUE)/sqrt(length(eggs_per_mate))
```

```{r em-v-fem-size, echo=FALSE, fig.cap="Scatterplot of the relationship between female size metrics and the number of eggs transferred.", fig.width=15, fig.height=5}
#create a subset of only the females who mated
mated_femSS <- fem_succSS[fem_succSS$MatingSuccess != 0, ]

#Plot total number of eggs transferred against the different metrics for females size
par(mfrow=c(1,3))
plot(mated_femSS$length,
     mated_femSS$totalEggs,
     xlab = "Length (mm)",
     ylab = "Number of eggs transferred",
     col = "darkorange",
     pch = 19)
abline(lm(mated_femSS$totalEggs ~ mated_femSS$length), lwd = 3, lty = 2)
plot(mated_femSS$depth_adj,
     mated_femSS$totalEggs,
     xlab = "Depth (mm)",
     ylab = "Number of eggs transferred",
     col = "cyan4",
     pch = 19)
abline(lm(mated_femSS$totalEggs ~ mated_femSS$depth_adj), lwd = 3, lty = 2)
plot(mated_femSS$svl,
     mated_femSS$totalEggs,
     xlab = "Snout-vent Length (mm)",
     ylab = "Number of eggs transferred",
     col = "purple",
     pch = 19)
abline(lm(mated_femSS$totalEggs ~ mated_femSS$svl), lwd = 3, lty = 2)

```

There also appears to be a correlation between female body size and the number of eggs transferred, especially in terms of depth and snout-vent length. Let's run some correlations tests to see what they say.

```{r eggs-fem-size-cor, echo=FALSE}
cor.test(mated_femSS$length, as.numeric(mated_femSS$totalEggs))
cor.test(mated_femSS$depth_adj, as.numeric(mated_femSS$totalEggs))
cor.test(mated_femSS$svl, as.numeric(mated_femSS$totalEggs))

```

There is no sig. correlation between length or svl and the number of eggs transferred but we do see a significantly positive relationship between depth and number of eggs transferred! It seems to be depth is a good indicator of fecundity for females. 

# Differences between mated individuals and unmated individuals
I want to now see if there are any significant differences in the sizes of individuals who mated vs individuals that didn't mate in males and females. I am going to be focusing on the same morphometrics outlined above.

## Males
### Visual Comparison
Before conducting any analyses, let's see if we can visually detect any differences between males who mated and unmated individuals.

```{r mat-status-morph-mal, echo=FALSE, fig.cap="_Six different morphometrics compared between males who sucessfully mated versus those that didn't. Orange represents unmated and blue represents mated males._", fig.height=10, fig.width=15, warning=FALSE}

par(mfrow=c(2,3))
boxplot(mal_succSS$length ~ as.factor(mal_succSS$MatingSuccess),
        xlab = "Mated",
        ylab = "Length (mm)",
        col= c("darkorange", "cyan4"))
boxplot(mal_succSS$depth_adj ~ as.factor(mal_succSS$MatingSuccess),
        xlab = "Mated",
        ylab = "Depth (mm)",
        col= c("darkorange", "cyan4"))
boxplot(mal_succSS$svl ~ as.factor(mal_succSS$MatingSuccess),
        xlab = "Mated",
        ylab = "Snout-vent Length (mm)",
        col= c("darkorange", "cyan4"))
boxplot(as.numeric(mal_succSS$bp_area) ~ as.factor(mal_succSS$MatingSuccess),
        xlab = "Mated",
        ylab = expression(paste("Brood Pouch Area (mm"^2*")")),
        col= c("darkorange", "cyan4"))
boxplot(mal_succSS$bp_length ~ as.factor(mal_succSS$MatingSuccess),
        xlab = "Mated",
        ylab = "Brood Pouch Length (mm)",
        col= c("darkorange", "cyan4"))

```

I don't notice many differences, however, it appears that males who mated may be slightly larger.

### Testing the difference
Let's now put some statistical power behind the difference in various morphometrics between mated and unmated individuals. I am first going to test the assumptions and then run the appropriate version of a t-test. 

```{r mat-stat-stat-mal, echo=FALSE, warning=FALSE}
##Test the assumptions
##Looking for equal variances
var.test(mal_succSS$length ~ as.factor(mal_succSS$MatingSuccess)) #equal
var.test(mal_succSS$depth_adj ~ as.factor(mal_succSS$MatingSuccess)) #equal
var.test(mal_succSS$svl ~ as.factor(mal_succSS$MatingSuccess)) #equal
var.test(mal_succSS$bp_area ~ as.factor(mal_succSS$MatingSuccess)) #equal
var.test(mal_succSS$bp_length ~ as.factor(mal_succSS$MatingSuccess)) #equal

##Looking for a normal distribution
normalTest(mal_succSS$length, method = "da") #normal
normalTest(mal_succSS$depth_adj, method = "da") #normal
normalTest(mal_succSS$svl, method = "da") #normal
normalTest(mal_succSS$bp_area, method = "da") #normal
normalTest(mal_succSS$bp_length, method = "da") #not normal

##Running the appropriate test
t.test(mal_succSS$length ~ as.factor(mal_succSS$MatingSuccess),
       var.equal = TRUE) #Not sig.
t.test(mal_succSS$depth_adj ~ as.factor(mal_succSS$MatingSuccess),
       var.equal = TRUE) #not sig.
t.test(mal_succSS$svl ~ as.factor(mal_succSS$MatingSuccess),
       var.equal = TRUE) #not sig.
t.test(mal_succSS$bp_area ~ as.factor(mal_succSS$MatingSuccess),
       var.equal = TRUE) #not sig.
wilcox.test(mal_succSS$bp_length ~ as.factor(mal_succSS$MatingSuccess)) #not sig.

```

There were no significant differences in terms of the sizes of males who mated versus those who did not. This may also be a result of the small sample size for unmated individuals.

Let's explore this a bit more and overlay the distribution of all males (mated and unmated) with the males who did mate and see how it varies for snout-vent length and depth.

```{r mated-unmated-hist, echo=FALSE, fig.cap= "_Overlay of the size range of males who mated on top of the size range of all males for snout-vent length (left) and torso depth (right)._", fig.width= 15, fig.height= 8}

par(mfrow = c(1,2))
hist(mal_succSS$svl, 
     xlab = "Snout-vent length (mm)", 
     ylab = "Number of male pipefish",
     col = "#beaed4", breaks = 15, main = "")

hist(mal_succSS$svl[mal_succSS$MatingSuccess != 0], 
     add = TRUE,
     col = "#8968CD", breaks = 15)

hist(mal_succSS$depth_adj, 
     xlab = "Torso depth (mm)", 
     ylab = "Number of male pipefish",
     col = "#beaed4", breaks = 15, main = "")

hist(mal_succSS$depth_adj[mal_succSS$MatingSuccess != 0], 
     add = TRUE,
     col = "#8968CD", breaks = 15)


legend("topright",
       legend = c("Mated Males", "All Males"),
       col = c("#8968CD", "#beaed4"),
       pch = 15)
```


## Females
Similarly, now let's see if we can identify any significant differences in the morphometrics of females who were able to obtain mates versus those who were unsuccessful.

### Visual Comparison

```{r mat-status-morph-fem, echo=FALSE, fig.cap="_Four different morphometrics compared between females who sucessfully mated versus those that didn't. Orange represents unmated and blue represents mated females._", fig.height=10, fig.width=10}

par(mfrow=c(1,3))
fem_succSS$mated <- as.factor(fem_succSS$mated)

boxplot(fem_succSS$length ~ fem_succSS$mated,
        xlab = "Mated",
        ylab = "Length (mm)",
        col= c("darkorange", "cyan4"))
boxplot(fem_succSS$depth_adj ~ fem_succSS$mated,
        xlab = "Mated",
        ylab = "Depth (mm)",
        col= c("darkorange", "cyan4"))
boxplot(fem_succSS$svl ~ fem_succSS$mated,
        xlab = "Mated",
        ylab = "Snout-vent Length (mm)",
        col= c("darkorange", "cyan4"))

```

For all categories it appears that females are larger in length, depth, and snout-vent length.

### Testing the difference
Let's now put some statistical power behind the difference in various morphometrics between mated and unmated individuals. 

```{r mat-stat-stat-fem, echo=FALSE, warning=FALSE}
#Test the assumptions
##Equal variances
var.test(fem_succSS$length ~ fem_succSS$mated) #equal
var.test(fem_succSS$depth_adj ~ fem_succSS$mated) #equal
var.test(fem_succSS$svl ~ fem_succSS$mated) #equal

##Normal Distributions
normalTest(fem_succSS$length, method = "da") #not normal
normalTest(fem_succSS$depth_adj, method = "da") #normal
normalTest(fem_succSS$svl, method = "da") #not normal

#Running the appropriate test
wilcox.test(fem_succSS$length ~ fem_succSS$mated) #sig.
t.test(fem_succSS$depth_adj ~ fem_succSS$mated, 
       var.equal = TRUE) #sig.
wilcox.test(fem_succSS$svl ~ fem_succSS$mated) #sig.

```

Females who mated are significantly larger in terms of standard length, torso depth, and snout-vent length.

Let's explore this a bit more and overlay the distribution of sizes in all females (mated and unmated) with the sizes of females who did mate and see how it varies.

```{r mated-unmated-hist-depth, echo=FALSE, fig.cap= "_Overlay of the standard length, torso depth, and snout-vent length of females who mated on top of the size range of all females._", fig.width= 15, fig.height= 10}

par(mfrow = c(1,3))
hist(fem_succSS$length, 
     xlab = "Standard Length (mm)", 
     ylab = "Number of female pipefish",
     col = "#7fc97f", breaks = 15, main = "")

hist(fem_succSS$length[fem_succSS$MatingSuccess != 0], 
     add = TRUE,
     col = "#548B54", breaks = 15)

hist(fem_succSS$depth_adj, 
     xlab = "Torso Depth (mm)", 
     ylab = "Number of female pipefish",
     col = "#7fc97f", breaks = 12, main = "")

hist(fem_succSS$depth_adj[fem_succSS$MatingSuccess != 0], 
     add = TRUE,
     col = "#548B54", breaks = 12)

hist(fem_succSS$svl, 
     xlab = "Snout-vent Length (mm)", 
     ylab = "Number of female pipefish",
     col = "#7fc97f", breaks = 20, main = "")

hist(fem_succSS$svl[fem_succSS$MatingSuccess != 0], 
     add = TRUE,
     col = "#548B54", breaks = 15)

legend("topright",
       legend = c("Mated Females", "All Females"),
       col = c("#548B54", "#7fc97f"),
       pch = 15)
```

There is still quite a spread in the sizes of females who mated, but in every case the female with the largest value secured a mate.

# Looking into the Opportunity for Selection in Males and Females
One of the benefits of using genetic parentage analysis is that we can now calculate the opportunity for selection and the opportunity for sexual selection in male and female pipefish. 

## Generating the total opportunity for selection ($I$) and the opportunity for sexual selection ($I_S$)
Because each trial provides an independent "population" (i.e., pipefish from one trial **cannot mate** with pipefish from another trial), I am going to calculate these metrics for each trial individually and then I will average it. With these I can then also generate 95% confidence intervals which I will investigate for indications of significance in two ways:

  - If the confidence intervals **DO NOT** cross 0 -> significant selection.
  - If the confidence intervals between the sexes **DO NOT** cross -> significantly different selection between the two sexes.

```{r opp-selection}
##FEMALES
#Create a dataframe to store the calculations of I and I_S in
fem_opp_selection <- data.frame(matrix(ncol = 3,
                                       nrow = 0))

colnames(fem_opp_selection) <- c("trial_num", "I", "I_s")

#Loop through the different trials and calculate I and I_S
for (trial in unique(fem_succSS$trial_num)) {
  
  #Subset the overall dataframe to work with an individual trial
  tmp <- fem_succSS[fem_succSS$trial_num == trial, ]
  
  #Calculate opportunity selection
  I <- var(tmp$NumDeveloped)/(mean(tmp$NumDeveloped)^2)
  
  I_s <- var(tmp$MatingSuccess)/(mean(tmp$MatingSuccess)^2)
  
  #Combining all of the selection values (Is) and save the output
  trial_num <- as.numeric(gsub("^(C)(\\d)", "\\2",
                               trial))
  selection <- cbind(trial_num, I, I_s)
  
  fem_opp_selection <- rbind(fem_opp_selection, selection)
  
}


##MALES
#Create a dataframe to store the calculations of I and I_S in
mal_opp_selection <- data.frame(matrix(ncol = 3,
                                       nrow = 0))

colnames(mal_opp_selection) <- c("trial_num", "I", "I_s")

#Loop through the different trials and calculate I and I_S
for (trial in unique(mal_succSS$trial_num)) {
  
  #Subset the overall dataframe to work with an individual trial
  tmp <- mal_succSS[mal_succSS$trial_num == trial, ]
  
  #Calculate opportunity selection
  I <- var(tmp$NumDeveloped)/(mean(tmp$NumDeveloped)^2)
  
  I_s <- var(tmp$MatingSuccess)/(mean(tmp$MatingSuccess)^2)
  
  #Combining all of the selection values (Is) and save the output
  trial_num <- as.numeric(gsub("^(C)(\\d)", "\\2",
                               trial))
  selection <- cbind(trial_num, I, I_s)
  
  mal_opp_selection <- rbind(mal_opp_selection, selection)
  
}

#Merge the selection coefficients from males and females into one dataset to 
#make life easier
fem_opp_selection$Sex <- "F"
mal_opp_selection$Sex <- "M"

opp_selection_all <- rbind(fem_opp_selection, mal_opp_selection)
```

Now that I have calculated the opportunity for selection and sexual selection, I want to generate my averages and 95% CI for both.

```{r opp-selection-CI}
#List the columns of interest
columns <- c("I", "I_s")

#Create a dataframe to store the final values in
opp_average <- data.frame(matrix(ncol = 4,
                                 nrow = 0))
colnames(opp_average) <- c("Average", "Interval", "Episode_sel", "Sex")

#Calculate the critical value
crit <- qt(p = 0.975, df = (nrow(fem_opp_selection) - 1))

for (j in 1:length(columns)) {
    
    col_name <- columns[[j]]
    
    #Calculate the means
    mean <- t(t(tapply(opp_selection_all[, colnames(opp_selection_all) == col_name], 
                       opp_selection_all$Sex, 
                       mean)))
    
    #Calculate standard error
    se <- t(t(tapply(opp_selection_all[, colnames(opp_selection_all) == col_name], 
                     opp_selection_all$Sex, 
                 function(x){
                   sqrt(var(x))/sqrt(length(x))
                 })))
    
    #Calculate the value that is added and subtracted from the mean
    int <- se*crit
    
    #Combine the data together
    episode <- as.data.frame(cbind(mean, int))
    colnames(episode) <- c("Average", "Interval")
    
    episode$Episode_sel <- col_name
    episode$Sex <- rownames(episode)
    
    rownames(episode) <- NULL
    
    opp_average <- rbind(opp_average, episode)
    
  }

```

Let's now explore some results:

```{r opp-selection-table, echo = FALSE, warning=FALSE}

#Create a new column
opp_average <- opp_average %>%
  mutate(Combined = paste0(round(Average , 2), 
                           " (", round(Average - Interval, 2), ", ",
                           round(Average + Interval, 2), ")"))

#Transform data into wide format
sum_opp <- opp_average %>% 
  select(Episode_sel, Sex, Combined) %>% 
  pivot_wider(names_from = Sex, values_from = Combined) 

#Display the table
kable(sum_opp, caption = "Average Opportunity of Selection (95% CI) for Males and Females")

```


```{r opp-selection-figure, echo=FALSE, fig.cap="_Average opportunity for selection and opportunity for sexual selection for male (purple) and female (green) S. scovelli. Errorbars represent the 95% confidence intervals around the mean_", fig.width=8, fig.height=6, warning=FALSE}

#Set parameters
sex_cols <- c("F" = "#7fc97f", "M" = "#beaed4" )
pd <- position_dodge(1)

ggplot(opp_average,
       aes(x = Episode_sel,
           y = Average,
           color = Sex,
           group = Sex)) +
  geom_errorbar(aes(ymin = Average - Interval,
                    ymax = Average + Interval),
                color = "black",
                width = 0.2,
                position = pd,
                linewidth = 1) +
  geom_point(position = pd,
             size = 4) +
  xlab("Episode of Selection") +
  ylab("Average Fitness Value") +
  scale_color_manual(values = sex_cols) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(legend.position = "bottom",
        legend.key=element_blank(),
        axis.title.x = element_text(size = 16,
                                    vjust = -0.5),
        axis.text.x = element_text(size = 16, family = "serif"),
        axis.title.y = element_text(size = 16,
                                    vjust = 2.5),
        axis.text.y = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 1)

```

We can see that for female _S. scovelli_ there is a significant opportunity for selection and opportunity for sexual selection, however, males only experience a significant opportunity for selection, but not sexual selection. Additionally, we can see a significant difference in the opportunity for sexual selection between males and females.

## Partitioning the Total Opportunity for Selection ($I$)
Once again for partitioning the opportunity for selection, I am going to calculate selection for the trials individually in males and females and then average across all trials to get the final values and the 95% CIs. For pre-mating processes I am focusing on mating success and then for post-mating processes I am looking at the total number of eggs transferred/received and the proportion of those eggs which developed (showing fertilization success).

```{r opp-select-episodes-females}

#Create a dataframe to store all of the intermediate values of fitness in
fem_succ_fitness <- data.frame(matrix(ncol = ncol(fem_succSS) + 9,
                                      nrow = 0))
colnames(fem_succ_fitness) <- c(colnames(fem_succSS),
                                "w1", "w1_squared",
                                "W2", "W2_bar", "w2",
                                "W3", "W3_bar", "w3", "i3")

#Create a dataframe to store the final calculations of I in
opp_selection_episodes_fem <- data.frame(matrix(ncol = 12,
                                            nrow = 0))
colnames(opp_selection_episodes_fem) <- c("trial_num", "I_1", "I_1per", "I_2", "I_2per", 
                                          "I_3", "I_3per", "I_12", "I_12per",
                                          "I", "Iper")

for (trial in unique(fem_succSS$trial_num)) {
  
  #Subset the overall dataframe to work with an individual trial
  tmp <- fem_succSS[fem_succSS$trial_num == trial, ]
  
  #Calculate the absolute pre-copulatory fitness (Eq. 14 Arnold & Wade 1984)
  #This is the same as the calculation of I_s
  tmp$w1 <- tmp$MatingSuccess/mean(tmp$MatingSuccess) #Relative mating success
  tmp$w1_squared <- (tmp$w1)^2
  
  I_1 <- var(tmp$w1) #Variance in relative mating success
  
  #Post-copulatory selection event 1 (Number of eggs transferred) (Eq. 15 Arnold & Wade 1984)
  tmp$W2 <- ifelse(tmp$MatingSuccess > 0,
                   tmp$totalEggs/tmp$MatingSuccess,
                   0) #Number of eggs per mate
  tmp$W2_bar <- tmp$W2 * (tmp$w1/nrow(tmp)) #Number of eggs per mate adjusted by the # of individuals with fitness W
  tmp$w2 <- tmp$W2/sum(tmp$W2_bar)
  
  I_2 <- (sum((tmp$w1 * (tmp$w2)^2))/nrow(tmp) - 1) * nrow(tmp)/(nrow(tmp) - 1)
  
  #Post-copulatory selection event 2 (Number of eggs developed) (Eq. 16 Arnold & Wade 1984)
  tmp$W3 <- ifelse(tmp$totalEggs > 0,
                   tmp$NumDeveloped/tmp$totalEggs,
                   0) #Proportion of transferred eggs that developed
  tmp$W3_bar <- tmp$W3 * ((tmp$totalEggs/mean(tmp$totalEggs))/nrow(tmp)) #Prop. of eggs developed adjusted by the # of individuals with fitness W
  tmp$w3 <- tmp$W3/sum(tmp$W3_bar)
  tmp$i3 <- ((tmp$totalEggs/mean(tmp$totalEggs))/nrow(tmp)) * ((tmp$w3 - 1)^2)
  
  I_3 <- sum(tmp$i3) * nrow(tmp)/(nrow(tmp) - 1)

  I_12 <- var(tmp$totalEggs)/(mean(tmp$totalEggs)^2)
  
  #Total opportunity for selection
  I <- var(tmp$NumDeveloped)/(mean(tmp$NumDeveloped)^2)
  
  #Calculating percentages for each selection event
  I_1per <- (I_1/I)*100
  I_2per <- (I_2/I)*100
  I_3per <- (I_3/I)*100
  I_12per <- (I_12/I)*100
  Iper <- (I/I)*100
  
  #Combining all of the selection values (Is) and saving the output
  trial_num <- as.numeric(gsub("^(C)(\\d)", "\\2",
                               trial))
  selection <- cbind(trial_num, I_1, I_1per, I_2, I_2per, I_3, I_3per,
                     I_12, I_12per, I, Iper)
  
  opp_selection_episodes_fem <- rbind(opp_selection_episodes_fem, selection)
  
  #Save the intermediate values
  fem_succ_fitness <- rbind(fem_succ_fitness, tmp)
}

#Exporting the data
#write.csv(fem_succ_fitness, "data/scovelli_int_I_fem.csv", row.names = FALSE)
```

```{r opp-select-episodes-males}
#Create a dataframe to store all of the intermediate values of fitness in
mal_succ_fitness <- data.frame(matrix(ncol = ncol(mal_succSS) + 9,
                                      nrow = 0))
colnames(mal_succ_fitness) <- c(colnames(mal_succSS),
                                "w1", "w1_squared",
                                "W2", "W2_bar", "w2",
                                "W3", "W3_bar", "w3", "i3")

#Create a dataframe to store the final calculations of I in
opp_selection_episodes_mal <- data.frame(matrix(ncol = 12,
                                            nrow = 0))
colnames(opp_selection_episodes_mal) <- c("trial_num", "I_1", "I_1per", "I_2", "I_2per", 
                                          "I_3", "I_3per", "I_12", "I_12per",
                                          "I", "Iper", "I_s")

for (trial in unique(mal_succSS$trial_num)) {
  
  #Subset the overall dataframe to work with an individual trial
  tmp <- mal_succSS[mal_succSS$trial_num == trial, ]
  
  #Calculate the absolute pre-copultory fitness (Eq. 14 Arnold & Wade 1984)
  tmp$w1 <- tmp$MatingSuccess/mean(tmp$MatingSuccess) #Relative mating success
  tmp$w1_squared <- (tmp$w1)^2
  
  I_1 <- var(tmp$w1) #Variance in relative mating success
  
  #Post-copulatory selection event 1 (Number of eggs transferred) (Eq. 15 Arnold & Wade 1984)
  tmp$W2 <- ifelse(tmp$MatingSuccess > 0,
                   tmp$totalEggs/tmp$MatingSuccess,
                   0) #Number of eggs per mate
  tmp$W2_bar <- tmp$W2 * (tmp$w1/nrow(tmp)) #Number of eggs per mate adjusted by the # of individuals with fitness W
  tmp$w2 <- tmp$W2/sum(tmp$W2_bar)
  
  I_2 <- (sum((tmp$w1 * (tmp$w2)^2))/nrow(tmp) - 1) * nrow(tmp)/(nrow(tmp) - 1)
  
  #Post-copulatory selection event 2 (Number of eggs developed) (Eq. 16 Arnold & Wade 1984)
  tmp$W3 <- ifelse(tmp$totalEggs > 0,
                   tmp$NumDeveloped/tmp$totalEggs,
                   0) #Proportion of transferred eggs that developed
  tmp$W3_bar <- tmp$W3 * ((tmp$totalEggs/mean(tmp$totalEggs))/nrow(tmp)) #Prop. of eggs developed adjusted by the # of individuals with fitness W
  tmp$w3 <- tmp$W3/sum(tmp$W3_bar)
  tmp$i3 <- ((tmp$totalEggs/mean(tmp$totalEggs))/nrow(tmp)) * ((tmp$w3 - 1)^2)
  
  I_3 <- sum(tmp$i3) * nrow(tmp)/(nrow(tmp) - 1)

  I_12 <- var(tmp$totalEggs)/(mean(tmp$totalEggs)^2)
  
  #Total opportunity for selection
  I <- var(tmp$NumDeveloped)/(mean(tmp$NumDeveloped)^2)

  #Calculating percentages for each selection event
  I_1per <- (I_1/I)*100
  I_2per <- (I_2/I)*100
  I_3per <- (I_3/I)*100
  I_12per <- (I_12/I)*100
  Iper <- (I/I)*100
  
  #Combining all of the selection values (Is) and saving the output
  trial_num <- as.numeric(gsub("^(C)(\\d)", "\\2",
                               trial))
  selection <- cbind(trial_num, I_1, I_1per, I_2, I_2per, I_3, I_3per,
                     I_12, I_12per, I, Iper)
  
  opp_selection_episodes_mal <- rbind(opp_selection_episodes_mal, selection)
  
  #Save the intermediate values
  mal_succ_fitness <- rbind(mal_succ_fitness, tmp)
}

#Exporting the data
#write.csv(mal_succ_fitness, "data/scovelli_int_I_mal.csv", row.names = FALSE)
```

```{r select-episodes-CI}
#Merge the selection coefficients from males and females into one dataset to 
#make life easier
opp_selection_episodes_fem$Sex <- "F"
opp_selection_episodes_mal$Sex <- "M"

opp_selection_episodes_all <- rbind(opp_selection_episodes_fem, opp_selection_episodes_mal)

#Exporting the data
#write.csv(opp_selection_episodes_all, "data/scovelli_opp_selection.csv", row.names = FALSE)

#List the columns of interest
columns <- c("I_1", "I_2", "I_12", "I_3","I")

#Create a dataframe to store the final values in
opp_episodes_average <- data.frame(matrix(ncol = 4,
                                    nrow = 0))
colnames(opp_episodes_average) <- c("Average", "Interval", 
                                    "Episode_sel", "Sex")

#Calculate the critical value
crit <- qt(p = 0.975, df = (nrow(opp_selection_episodes_fem) - 1))

for (j in 1:length(columns)) {
    
    col_name <- columns[[j]]
    
    #Calculate the means
    mean <- t(t(tapply(opp_selection_episodes_all[, colnames(opp_selection_episodes_all) 
                                                  == col_name], 
                       opp_selection_episodes_all$Sex, 
                       mean)))
    
    #Calculate standard error
    se <- t(t(tapply(opp_selection_episodes_all[, colnames(opp_selection_episodes_all) 
                                                == col_name], 
                     opp_selection_episodes_all$Sex, 
                 function(x){
                   sqrt(var(x))/sqrt(length(x))
                 })))
    
    #Calculate the value that is added and subtracted from the mean
    int <- se*crit
    
    #Combine the data together
    episode <- as.data.frame(cbind(mean, int))
    colnames(episode) <- c("Average", "Interval")
    
    episode$Episode_sel <- col_name
    episode$Sex <- rownames(episode)
    
    rownames(episode) <- NULL
    
    opp_episodes_average <- rbind(opp_episodes_average, episode)
    
  }


```

Let's now explore some results:

```{r opp-select-episodes-table, echo = FALSE, warning=FALSE}

#Create a new column
opp_episodes_average <- opp_episodes_average %>%
  mutate(Combined = paste0(round(Average , 3), 
                           " (", round(Average - Interval, 2), ", ",
                           round(Average + Interval, 2), ")"))

#Transform data into wide format
sum_table <- opp_episodes_average %>% 
  select(Episode_sel, Sex, Combined) %>% 
  pivot_wider(names_from = Sex, values_from = Combined) 

#Display the table
kable(sum_table, caption = "Average Episode of Selection (95% CI) for Males and Females")

```


```{r opp-select-episodes-figure, echo=FALSE, fig.cap="_Average opportunity for selection for the different episodes for male (purple) and female (green) S. scovelli. Errorbars represent the 95% confidence intervals around the mean_", fig.width=8, fig.height=6, warning=FALSE}

#Set parameters

ggplot(opp_episodes_average,
       aes(x = Episode_sel,
           y = Average,
           color = Sex,
           group = Sex)) +
  geom_errorbar(aes(ymin = Average - Interval,
                    ymax = Average + Interval),
                color = "black",
                width = 0.2,
                position = pd,
                linewidth = 1) +
  geom_point(position = pd,
             size = 4) +
  xlab("Episode of Selection") +
  ylab("Average Fitness Value") +
  scale_color_manual(values = sex_cols) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(legend.position = "bottom",
        legend.key=element_blank(),
        axis.title.x = element_text(size = 16,
                                    vjust = -0.5),
        axis.text.x = element_text(size = 16, family = "serif"),
        axis.title.y = element_text(size = 16,
                                    vjust = 2.5),
        axis.text.y = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  geom_hline(yintercept = 0, linetype = "dashed", size = 1)

```

From the table and the plot we can see that once again there are some significant differences in the selection between males and females. 

Let's now look more into the percentage of the overall opportunity for selection made up for by each individual episode of selection:

```{r calc-opp-percents}

sexes <- c("M", "F")

#Create a dataframe to store the final values in
opp_percents <- data.frame(matrix(ncol = 3,
                                 nrow = 0))
colnames(opp_percents) <- c("Percent", "Episode_sel", "Sex")

#Calculate the percentage for each episode in the two sexes
for (sex in sexes) {
    
  #subset dataset based on sex
  tmp_sex <- opp_episodes_average[opp_episodes_average$Sex == sex, ]
  
  #Pull out the overall opp. for selection value
  I_average <- tmp_sex$Average[tmp_sex$Episode_sel == "I"]
  
  #Calculate what percentage of the I is represented by each episode
  percents <- as.data.frame(t(t(apply(tmp_sex, 1, function(x){
      
      (as.numeric(x[1])/I_average)*100
      
    }))))
    
    colnames(percents) <- "Percent"
    percents$Episode_sel <- tmp_sex$Episode_sel
    percents$Sex <- sex
    
    rownames(percents) <- NULL
    
    opp_percents <- rbind(opp_percents, percents)
    
  }

```

```{r generate-figa-opp-selection, echo=FALSE, warning=FALSE, fig.cap="_The proportion of the total opportunity for selection that is represented by each episode of selection for males and females._"}

selection_colors <- c("I_1" = "#9FBCAA", "I_2" = "#F0B67F", "I_3" = "#CD96CD", "I_12" = "#6CA6CD")

ggplot(opp_percents[!(opp_percents$Episode_sel %in% c("I", "I_12")),], 
       aes(x = Sex, 
           y = Percent, 
           fill = Episode_sel)) +
  geom_bar(position = "stack", stat = "identity") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  scale_fill_manual(values = selection_colors) +
   theme(legend.position = "right",
        legend.key=element_blank(),
        axis.title.x = element_text(size = 16,
                                    vjust = -0.5),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 16,
                                    vjust = 2.5),
        axis.text.y = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12, family = "serif"))

```

Matching the previous plots, most of the opportunity for selection in _Syngnathus scovelli_ females can be attributed to variance in mating success ($I_1$) rather than variance in eggs transferred/received ($I_2$) or variance in the proportion of eggs developed ($I_3$). However, for males it appears that post-copulatory selection may be more important. This may be a result of males only having the capacity to mate once, meaning there is less variation in the mating success (i.e, they either mate or they don't).

# Mate success versus Reproductive success (Bateman Gradient, $\beta_{SS}$)
To calculate $\beta_{SS}$ we use _relative_ measures of fitness:
($\frac{indvidual's fitness}{mean(fitness)}$)

I am going to generate the measurements of relative fitness once again for the trials individually rather than across all of the trials.

```{r calc-relative-fitness}

#Calculating relative fitness as a metric for reproductive success
#Create a dataframe to store all of the calculations of relative fitness in
fem_bateman <- data.frame(matrix(ncol = 3,
                                 nrow = 0))
colnames(fem_bateman) <- c("trial", "MatingSuccess","rel_repo_fitness")

#Loop through each trial to calculate relative fitness
for (trial in unique(fem_succSS$trial_num)) {
  
  #Subset the overall dataframe to work with an individual trial
  tmp <- fem_succSS[fem_succSS$trial_num == trial, ]
  
  #Calculate relative fitness
  rel_repo_fitness <- tmp$totalEggs/mean(tmp$totalEggs)
  
  #Calculte mating fitness
  rel_mate_succuess <- tmp$MatingSuccess/mean(tmp$MatingSuccess)
  
  #Column-bind the trial #, Mating success, and calculated rel. fitness
  trial_num <- as.numeric(gsub("^(C)(\\d)", "\\2",
                               trial))
  fitness <- cbind("trial" = rep(trial_num, nrow(tmp)), 
                   "MatingSuccess" = rel_mate_succuess, 
                   rel_repo_fitness)
  
  #Add this chunk of data to the dataframe we created
  fem_bateman <- rbind(fem_bateman, fitness)
}


#Repeat process for the Male mating data
mal_bateman <- data.frame(matrix(ncol = 3,
                                 nrow = 0))
colnames(mal_bateman) <- c("trial", "MatingSuccess","rel_repo_fitness")

for (trial in unique(mal_succSS$trial_num)) {
  
  #Subset the overall dataframe to work with an individual trial
  tmp <- mal_succSS[mal_succSS$trial_num == trial, ]
  
  #Calculate relative fitness
  rel_repo_fitness <- tmp$totalEggs/mean(tmp$totalEggs)
  
  #Calculate mating fitness
  rel_mate_succuess <- tmp$MatingSuccess/mean(tmp$MatingSuccess)
  
  #Column-bind the trial #, Mating success, and calculated rel. fitness
  trial_num <- as.numeric(gsub("^(C)(\\d)", "\\2",
                               trial))
  fitness <- cbind("trial" = rep(trial_num, nrow(tmp)), 
                   "MatingSuccess" = rel_mate_succuess, 
                   rel_repo_fitness)
  
  #Add this chunk of data to the dataframe we created
  mal_bateman <- rbind(mal_bateman, fitness)
}

```

Once we have the measures of relative fitness we can use them to run the weighted least-squares regression for males and females separately.

```{r calc-bateman-gradient}

#Generating Bateman's gradient
#Define the model
fem_model <- lm(fem_bateman$rel_repo_fitness ~ fem_bateman$MatingSuccess)
mal_model <- lm(mal_bateman$rel_repo_fitness ~ mal_bateman$MatingSuccess)

#define weights to use
wt_fem <- 1 / lm(abs(fem_model$residuals) ~ fem_model$fitted.values)$fitted.values^2
wt_mal <- 1 / lm(abs(mal_model$residuals) ~ mal_model$fitted.values)$fitted.values^2

#perform weighted least squares regression
wls_model_fem <- lm(fem_bateman$rel_repo_fitness ~ fem_bateman$MatingSuccess,
                    weights=wt_fem)
wls_model_mal <- lm(mal_bateman$rel_repo_fitness ~ mal_bateman$MatingSuccess,
                    weights=wt_mal)

#Investigate the results
summary(wls_model_fem) #significant
summary(wls_model_mal) #significant

```

For both males and females there is a significant slope, meaning both sexes see increases in fitness with each additional mating. I am interested to run the model with the two datasets combined to see if there is an interaction of sex.

```{r calc-bateman-all}

#Combine the two datasets
fem_bateman$Sex <- "F"
mal_bateman$Sex <- "M"

all_bateman <- rbind(fem_bateman, mal_bateman)

#Running a weighted least squares regression between MS and Sex
MS_sex_model <- lm(all_bateman$rel_repo_fitness ~
                     all_bateman$MatingSuccess*all_bateman$Sex)

wt_all <- 1 / lm(abs(MS_sex_model$residuals) ~
                   MS_sex_model$fitted.values)$fitted.values^2

wls_MS_sex_model <- lm(all_bateman$rel_repo_fitness ~ 
                         all_bateman$MatingSuccess*all_bateman$Sex, weights = wt_all)

summary(wls_MS_sex_model)

```

When we combine all of the data together we can see two additional things:

  1. There is no significant interactions effect of Mating success and sex
  2. The slopes for males and females are not significantly different from each other
  
Let's visually look at this pattern now:

```{r plot-bateman3, echo=FALSE, fig.cap="_Relationship between reproductive success and mating success for male (purple) and female (green) _Syngnathus scovelli_. Reproductive success is shown as relative fitness (i.e. number of offspring produced divided by the mean number of offspring produced). Bateman's gradient is shown as the weighted least-squares regression line (dashed) for males and females._", fig.height=5}

#Plotting the relationship between the mating and reproductive success
plot(fem_bateman$MatingSuccess,
     fem_bateman$rel_repo_fitness,
     xlab = "Number of Mates (mating success)",
     ylab = "Relative Fitness (reproductive success)",
     pch = 19,
     col = "#7fc97f75")
points(mal_bateman$MatingSuccess,
       mal_bateman$rel_repo_fitness,
       xlab = "Number of Mates (mating success)",
       ylab = "Relative Fitness (reproductive success)",
       pch = 19,
       col = "#beaed475")
abline(wls_model_fem, col = "#7fc97f", lwd = 3, lty = 2)
abline(wls_model_mal, col = "#beaed4", lwd = 3, lty = 2)

legend("topleft",
       legend = c("Male", "Female"),
       pch = 16,
       cex = 1.25,
       col = c("#beaed4", "#7fc97f"))

#write.csv(fem_bateman, file = "data/SS_fem_bateman.csv", row.names = FALSE)
#write.csv(mal_bateman, file = "data/SS_mal_bateman.csv", row.names = FALSE)
```

The plot confirms the results from the two models. We can see that there is a steep slope for both males and females, however, the two lines are not different.

## Investigating the impact of "zeros" on the Bateman Gradient
It has been shown previously that the inclusion or exclusion of individuals who were unable to achieve a mate. When the non-mated individuals are included, the gradient is influenced by the relative fitness gain that is experienced from gaining a mate (i.e, moving from 0 to 1). If we do not include the non-mated individuals, we can more clearly look at the fitness increase associated with obtaining more than one mate.

For _S. scovelli_, because males only mate once it does not make sense to re-do the analysis without the non-mated individuals. Therefore, I will only be focusing on the females.

### Removing the zeros from the plot
The first way I am addressing this is by using the same datasets as before and just excluding the 0's from the plot and the model:

```{r calc-bateman-fem-nozero1}

#Generating Bateman's gradient
#Define the model
fem_model2 <- lm(fem_bateman$rel_repo_fitness[fem_bateman$MatingSuccess != 0] ~
                   fem_bateman$MatingSuccess[fem_bateman$MatingSuccess != 0])

#define weights to use
wt_fem2 <- 1 / lm(abs(fem_model2$residuals) ~
                    fem_model2$fitted.values)$fitted.values^2

#perform weighted least squares regression
wls_model_fem2 <- lm(fem_bateman$rel_repo_fitness[fem_bateman$MatingSuccess != 0] ~
                       fem_bateman$MatingSuccess[fem_bateman$MatingSuccess != 0],
                    weights=wt_fem2)

#Investigate the results
summary(wls_model_fem2) #significant

```

```{r plot-bateman2, echo=FALSE, fig.cap="_Relationship between reproductive success and mating success for female _Syngnathus scovelli_ who achieved at least one mate. Reproductive success is shown as relative fitness (i.e. number of offspring produced divided by the mean number of offspring produced). Bateman's gradient is shown as the weighted least-squares regression line (dashed)._", fig.height=5}

#Plotting the relationship between the mating and reproductive success
plot(fem_bateman$MatingSuccess[fem_bateman$MatingSuccess != 0],
     fem_bateman$rel_repo_fitness[fem_bateman$MatingSuccess != 0],
     xlab = "Number of Mates (mating success)",
     ylab = "Relative Fitness (reproductive success)",
     pch = 19,
     col = "#7fc97f75")

abline(wls_model_fem2, col = "#7fc97f", lwd = 3, lty = 2)

```

When we exclude the non-mated individuals from the plot and the model, we can see the results do not change for females, the slope is still significant.

### Removing the zeros from the calculation of relative fitness
The other way to approach this is rather than just eliminating the zeros from the scatter plot, I can remove the zeros in the calculation of relative fitness. 

```{r calc-relative-fitness2}

#Calculating relative fitness as a metric for reproductive success
#Create a dataframe to store all of the calculations of relative fitness in
fem_bateman_nozero <- data.frame(matrix(ncol = 3,
                                 nrow = 0))
colnames(fem_bateman_nozero) <- c("trial", "MatingSuccess","rel_repo_fitness")

#Loop through each trial to calculate relative fitness
for (trial in unique(fem_succSS$trial_num)) {
  
  #Subset the overall dataframe to work with an individual trial
  tmp <- fem_succSS[fem_succSS$trial_num == trial, ]
  
  #Calculate relative fitness
  rel_repo_fitness <- tmp$totalEggs[tmp$MatingSuccess != 0]/
    mean(tmp$totalEggs[tmp$MatingSuccess != 0])
  
  #Calculte mating fitness
  rel_mate_succuess <- tmp$MatingSuccess[tmp$MatingSuccess != 0]/
    mean(tmp$MatingSuccess[tmp$MatingSuccess != 0])
  
  #Column-bind the trial #, Mating success, and calculated rel. fitness
  trial_num <- as.numeric(gsub("^(C)(\\d)", "\\2",
                               trial))
  fitness <- cbind("trial" = rep(trial_num, nrow(tmp[tmp$MatingSuccess != 0,])), 
                   "MatingSuccess" = rel_mate_succuess, 
                   rel_repo_fitness)
  
  #Add this chunk of data to the dataframe we created
  fem_bateman_nozero <- rbind(fem_bateman_nozero, fitness)
}

```

Once we have the measures of relative fitness we can use them to run the weighted least-squares regression for males and females separately.

```{r calc-bateman-gradient2}

#Generating Bateman's gradient
#Define the model
fem_model3 <- lm(fem_bateman_nozero$rel_repo_fitness ~ 
                  fem_bateman_nozero$MatingSuccess)

#define weights to use
wt_fem3 <- 1 / lm(abs(fem_model3$residuals) ~
                    fem_model3$fitted.values)$fitted.values^2

#perform weighted least squares regression
wls_model_fem3 <- lm(fem_bateman_nozero$rel_repo_fitness ~
                      fem_bateman_nozero$MatingSuccess,
                    weights=wt_fem3)


#Investigate the results
summary(wls_model_fem3) #significant

```

```{r plot-bateman, echo=FALSE, fig.cap="_Relationship between reproductive success and mating success for female _Syngnathus scovelli_ who achieved at least one mate. Reproductive success is shown as relative fitness (i.e. number of offspring produced divided by the mean number of offspring produced). Relative fitness is calculated without the individuals who did not mate. Bateman's gradient is shown as the weighted least-squares regression line (dashed)._", fig.height=5}

#Plotting the relationship between the mating and reproductive success
plot(fem_bateman_nozero$MatingSuccess,
     fem_bateman_nozero$rel_repo_fitness,
     xlab = "Number of Mates (mating success)",
     ylab = "Relative Fitness (reproductive success)",
     pch = 19,
     col = "#7fc97f75")

abline(wls_model_fem3, col = "#7fc97f", lwd = 3, lty = 2)

```

With this way of excluding the individuals who did not mate, there is still a significant increase in relative fitness with each additional mating.

# Investing selection differentials on snout-vent-length ($s$ and $s'$)
A selection differential is the covariance between a trait and relative fitness and is often thought of as the difference in a trait mean before and after selection. 

Just as I did for the decomposition of the opportunity for selection, I am going to decompose the total selection differentials into the different pre- and post-mating episodes.

To keep things consistent with the analysis I conducted on _S. floridae_, I will once again be investigating snout-vent length as the trait of interest, even though standard length and torso depth was also significantly different between males and females.
  
```{r selection-diff-females}
#Create a dataframe to store all of the intermediate values of fitness in
fem_succ_select_diff <- data.frame(matrix(ncol = ncol(fem_succSS) + 6,
                                          nrow = 0))
colnames(fem_succ_select_diff) <- c(colnames(fem_succSS),
                                    "fit1", "eggs_per_mate","fit2", 
                                    "prop_dev", "fit3", "StdLength")

#Create a dataframe to store the final calculations of I in
select_diff_fem <- data.frame(matrix(ncol = 11,
                                     nrow = 0))
colnames(select_diff_fem) <- c("trial", "s1", "s2", "s3", "s12", "s123",
                               "s1_prime", "s2_prime", "s3_prime", 
                               "s12_prime", "s123_prime")

for (trial in unique(fem_succSS$trial_num)) {
  
  #Subset the overall dataframe to work with an individual trial
  tmp <- fem_succSS[fem_succSS$trial_num == trial, ]
  
  #Calculate fitness relating to pre-cop. selection (#matings)
  tmp$fit1 <- tmp$MatingSuccess/mean(tmp$MatingSuccess) #Relative mating success

  #Calculate fitness relating to post-mating selection (#eggs transferred)
  tmp$eggs_per_mate <- tmp$totalEggs/tmp$MatingSuccess
  ##If mating success = 0, eggs_per_mate = NA and it not included in the calculation
  ##of the relative fitness moving forward
  tmp$fit2 <- ifelse(tmp$MatingSuccess > 0,
                     tmp$eggs_per_mate/mean(tmp$eggs_per_mate, na.rm = TRUE),
                     0) #Relative eggs transferred

  #Calculate fitness relating to post-mating selection (eggs that developed)
  tmp$prop_dev <- (tmp$NumDeveloped/tmp$MatingSuccess)/tmp$eggs_per_mate
  tmp$fit3 <- ifelse(tmp$MatingSuccess > 0,
                     tmp$prop_dev/mean(tmp$prop_dev, na.rm = TRUE),
                     0)
  
  #Standardizing the trait value to have a mean of 0 and sd of unity
  tmp$StdLength <- (tmp$svl - mean(tmp$svl))/sd(tmp$svl)
  
  #Calculating the absolute selection differentials (s)
  s1 <- cov(tmp$svl, tmp$fit1)
  s12 <- cov(tmp$svl, tmp$fit2)
  s123 <- cov(tmp$svl, tmp$fit3)
  s2 <- s12 - s1
  s3 <- s123 - s12
  
  #Calculating the standardized selection differentials (s')
  s1_prime <- cov(tmp$StdLength, tmp$fit1)
  s12_prime <- cov(tmp$StdLength, tmp$fit2)
  s123_prime <- cov(tmp$StdLength, tmp$fit3)
  s2_prime <- s12_prime - s1_prime
  s3_prime <- s123_prime - s12_prime
  
  #Combining all of the selection differentials (s, s') and saving the output
  trial_num <- as.numeric(gsub("^(C)(\\d)", "\\2",
                               trial))
  selection <- cbind(trial_num, s1, s2, s3, s12, s123, 
                     s1_prime, s2_prime, s3_prime, s12_prime, s123_prime)
  
  select_diff_fem <- rbind(select_diff_fem, selection)
  
  #Save the intermediate values
  fem_succ_select_diff <- rbind(fem_succ_select_diff, tmp)
}

#Exporting the data
write.csv(fem_succ_select_diff, "data/scovelli_int_diff_fem.csv", row.names = FALSE)
```

```{r selection-diff-males}
#Create a dataframe to store all of the intermediate values of fitness in
mal_succ_select_diff <- data.frame(matrix(ncol = ncol(mal_succSS) + 6,
                                          nrow = 0))
colnames(mal_succ_select_diff) <- c(colnames(mal_succSS),
                                    "fit1", "eggs_per_mate","fit2", "prop_dev", 
                                    "fit3", "StdLength")

#Create a dataframe to store the final calculations of I in
select_diff_mal <- data.frame(matrix(ncol = 11,
                                     nrow = 0))
colnames(select_diff_mal) <- c("trial", "s1", "s2", "s3", "s12", "s123",
                               "s1_prime", "s2_prime", "s3_prime", 
                               "s12_prime", "s123_prime")

for (trial in unique(mal_succSS$trial_num)) {
  
  #Subset the overall dataframe to work with an individual trial
  tmp <- mal_succSS[mal_succSS$trial_num == trial, ]
  
  #Calculate fitness relating to pre-cop. selection (#matings)
  tmp$fit1 <- tmp$MatingSuccess/mean(tmp$MatingSuccess) #Relative mating success

  #Calculate fitness relating to post-mating selection (#eggs transferred)
  tmp$eggs_per_mate <- tmp$totalEggs/tmp$MatingSuccess
  tmp$fit2 <- ifelse(tmp$MatingSuccess > 0,
                     tmp$eggs_per_mate/mean(tmp$eggs_per_mate, na.rm = TRUE),
                     0) #Relative eggs transferred

  #Calculate fitness relating to post-mating selection (eggs that developed)
  tmp$prop_dev <- (tmp$NumDeveloped/tmp$MatingSuccess)/tmp$eggs_per_mate
  tmp$fit3 <- ifelse(tmp$MatingSuccess > 0,
                     tmp$prop_dev/mean(tmp$prop_dev, na.rm = TRUE),
                     0)
  
  #Standardizing the trait value to have a mean of 0 and sd of unity
  tmp$StdLength <- (tmp$svl - mean(tmp$svl))/sd(tmp$svl)
  
  #Calculating the absolute selection differentials (s)
  s1 <- cov(tmp$svl, tmp$fit1)
  s12 <- cov(tmp$svl, tmp$fit2)
  s123 <- cov(tmp$svl, tmp$fit3)
  s2 <- s12 - s1
  s3 <- s123 - s12
  
  #Calculating the standardized selection differentials (s')
  s1_prime <- cov(tmp$StdLength, tmp$fit1)
  s12_prime <- cov(tmp$StdLength, tmp$fit2)
  s123_prime <- cov(tmp$StdLength, tmp$fit3)
  s2_prime <- s12_prime - s1_prime
  s3_prime <- s123_prime - s12_prime
  
  #Combining all of the selection differentials (s, s') and saving the output
  trial_num <- as.numeric(gsub("^(C)(\\d)", "\\2",
                               trial))
  selection <- cbind(trial_num, s1, s2, s3, s12, s123, 
                     s1_prime, s2_prime, s3_prime, s12_prime, s123_prime)
  
  select_diff_mal <- rbind(select_diff_mal, selection)
  
  #Save the intermediate values
  mal_succ_select_diff <- rbind(mal_succ_select_diff, tmp)
}

#Exporting the data
write.csv(mal_succ_select_diff, "data/fuscus_int_diff_mal.csv", row.names = FALSE)
```

```{r selection-diff-CI}
#Merge the male and female datasets together
select_diff_fem$Sex <- "F"
select_diff_mal$Sex <- "M"

select_diff_all <- rbind(select_diff_fem, select_diff_mal)

write.csv(select_diff_all, "data/scovelli_select_diff.csv", row.names = FALSE)

#List the columns of interest
columns <- c("s1", "s2", "s3", "s123",
             "s1_prime", "s2_prime", "s3_prime", "s123_prime")

#Create a dataframe to store the final values in
sd_average <- data.frame(matrix(ncol = 4,
                                nrow = 0))
colnames(sd_average) <- c("Average", "Interval", "Select_diff", "Sex")

#Calculate the critical value
crit <- qt(p = 0.975, df = (nrow(select_diff_fem) - 1))

#Calculating the averages and confidence intervals for each species and 
#selection differential
for (j in 1:length(columns)) {
    
    col_name <- columns[[j]]
    
    #Calculate the means
    mean <- t(t(tapply(select_diff_all[, colnames(select_diff_all) == col_name], 
                       select_diff_all$Sex, mean)))
    
    #Calculate standard error
    se <- t(t(tapply(select_diff_all[, colnames(select_diff_all) == col_name], 
                     select_diff_all$Sex, 
                 function(x){
                   
                   sqrt(var(x))/sqrt(length(x))
                   
                 })))
    
    #Calculate the value that is added and subtracted from the mean
    int <- se*crit
    
    #Combine the data together
    episode <- as.data.frame(cbind(mean, int))
    colnames(episode) <- c("Average", "Interval")
    
    episode$Select_diff <- col_name
    episode$Sex <- rownames(episode)
    
    rownames(episode) <- NULL
    
    sd_average <- rbind(sd_average, episode)
    
}


```

Now that we have the average select diff for males and females alongside the 95% confidence intervals we can visualize some results:

```{r select-diff-table, echo = FALSE, warning=FALSE}

#Create a new column
sd_average <- sd_average %>%
  mutate(Combined = paste0(round(Average , 2), 
                           " (", round(Average - Interval, 2), ", ",
                           round(Average + Interval, 2), ")"))

#Transform data into wide format
sum_table <- sd_average %>% 
  select(Select_diff, Sex, Combined) %>% 
  pivot_wider(names_from = Sex, values_from = Combined) 

#Display the table
kable(sum_table, caption = "Average Selection Differentials (95% CI) for Males and Females")

```

```{r generate-fig-select-diff, echo=FALSE, fig.width=15, fig.height=8, fig.cap= "_Absolute (left) and standardized (right) selection differentials for male (purple) and female (green) S. fuscus. Error bars represent the 95% confidence intervals around the mean._"}
pd <- position_dodge(1)

select_diff_average <- ggplot(sd_average[sd_average$Select_diff %in% 
                                           c("s1", "s2", "s3", "s123"),],
       aes(x = Select_diff,
           y = Average,
           color = Sex,
           group = Sex)) +
  geom_errorbar(aes(ymin = Average - Interval,
                    ymax = Average + Interval),
                color = "black",
                width = 0.8,
                position = pd,
                linewidth = 1) +
  geom_point(position = pd,
             size = 4) +
  xlab("Absolute Selection Differential") +
  ylab("Average value across trials") +
  scale_color_manual(values = sex_cols) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(legend.position = "bottom",
        legend.key=element_blank(),
        axis.title.x = element_text(size = 16,
                                    vjust = -0.5),
        axis.text.x = element_text(size = 12, family = "serif"),
        axis.title.y = element_text(size = 16,
                                    vjust = 2.5),
        axis.text.y = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 1)

select_diff2_average <- ggplot(sd_average[sd_average$Select_diff %in% 
                                           c("s1_prime", "s2_prime", 
                                             "s3_prime", "s123_prime"),],
       aes(x = Select_diff,
           y = Average,
           color = Sex,
           group = Sex)) +
  geom_errorbar(aes(ymin = Average - Interval,
                    ymax = Average + Interval),
                color = "black",
                width = 0.8,
                position = pd,
                linewidth = 1) +
  geom_point(position = pd,
             size = 4) +
  xlab("Absolute Selection Differential") +
  ylab("Average value across trials") +
  scale_color_manual(values = sex_cols) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(legend.position = "bottom",
        legend.key=element_blank(),
        axis.title.x = element_text(size = 16,
                                    vjust = -0.5),
        axis.text.x = element_text(size = 12, family = "serif"),
        axis.title.y = element_text(size = 16,
                                    vjust = 2.5),
        axis.text.y = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 1)

plot_grid(select_diff_average, select_diff2_average,
          ncol = 2)

```

We can see from these results that males and females are experiencing largely similar selection on snout-vent length except for in the first episode of post-mating selection (number of eggs transferred). 

Overall, it appears larger size in svl is more beneficial. However, while we can see these patterns, there is only significant selection on standardized snout-vent length (all the 95% CI cross 0) for females in pre-mating selection ($s_1$).

## Looking into the Maximum Sexual Selection Differential
We can calculate $s'_{max}$:

$$
s'_{max}\ =\ \beta_{SS}\sqrt{I_S}
$$

One major benefit of this is that both the variation in mating success and the Bateman gradient are taken into effect. Additionally, $s'_{max}$ places an upper bound on the strength of sexual selection during a sexual selection episode.

Let's now generate $s'_{max}$ for male and female _S. scovelli_.

```{r select-diff-max}
#Pull out the Bateman gradient (slopes of the models)
bateman_fem <- coefficients(wls_model_fem)[2]
bateman_mal <- coefficients(wls_model_mal)[2]

#Pull out the Average opp sexual selection for males and females
opp_ss_fem <- opp_average$Average[opp_average$Sex == "F" & 
                                    opp_average$Episode_sel == "I_s"]
opp_ss_mal <- opp_average$Average[opp_average$Sex == "M" & 
                                    opp_average$Episode_sel == "I_s"]

#Calculate select diff max
select_diff_max_fem <- bateman_fem * sqrt(opp_ss_fem)
select_diff_max_mal <- bateman_mal * sqrt(opp_ss_mal)

```

For females, the $s'_{max}$ is `r (select_diff_max_fem)`, while $s'$ is `r sd_average$Average[sd_average$Sex == "F" & sd_average$Select_diff == "s123_prime"]`, which is about 1/12 of the max selection that could be experienced. For males, $s'_{max}$ is `r (select_diff_max_mal)` and $s'$ is `r sd_average$Average[sd_average$Sex == "M" & sd_average$Select_diff == "s123_prime"]` which is only about 1/3 of the max selection that could be experienced.

# Visualizing post-copulatory selection
As a way to visualize selection acting AFTER the mating event (post-copulatory selection) I am plotting the proportion of eggs that survived against mating success. Hopefully this will tell us if acquiring more mates is having any affect on the ability for the eggs to develop.

I am going to plot this relationship only looking into the individuals who mated.

```{r surv-v-matings, echo=FALSE, fig.cap="_Plotting the relationship between the proportion of eggs that developed and the number of mates aquired for both males (purple) and females (green)._", fig.width=12, fig.height=8}

mal_succSS$prop_surviving <- ifelse(mal_succSS$totalEggs == 0, 0,
                                mal_succSS$NumDeveloped/mal_succSS$totalEggs)
fem_succSS$prop_surviving <- ifelse(fem_succSS$totalEggs == 0, 0,
                                  fem_succSS$NumDeveloped/fem_succSS$totalEggs)


plot(fem_succSS$MatingSuccess[fem_succSS$MatingSuccess != 0],
     fem_succSS$prop_surviving[fem_succSS$totalEggs != 0],
     xlab = "Number of Mates (mating success)",
     ylab = "Proportion of Surviving Offpring",
     pch = 19,
     col = "#7fc97f75",
     ylim = c(0, 1))
points(mal_succSS$MatingSuccess[mal_succSS$MatingSuccess != 0],
       mal_succSS$prop_surviving[mal_succSS$totalEggs != 0],
       xlab = "Number of Mates (mating success)",
       ylab = "Proportion of Surviving Offpring",
       pch = 19,
       col = "#beaed475")
abline(lm(fem_succSS$prop_surviving[fem_succSS$totalEggs != 0] ~
            fem_succSS$MatingSuccess[fem_succSS$MatingSuccess != 0]),
       col = "#7fc97f", lwd = 3, lty = 2)




```

There may be a correlation for females, let's investigate further:

```{r surv-mate-cor}

cor.test(fem_succSS$MatingSuccess[fem_succSS$MatingSuccess != 0],
         fem_succSS$prop_surviving[fem_succSS$totalEggs != 0])

```

There is a significantly negative correlation between mating success and reproductive success for female _S. scovelli_.

